A cautionary note on the use of Ornstein Uhlenbeck models in macroevolutionary studies

### 2015-06-23 Manuscript/Revision/response.tex

\documentclass[11pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[colorlinks = true, urlcolor = blue]{hyperref} %hyperlinks
\usepackage[osf]{mathpazo}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\signature{Natalie Cooper \\ Gavin Thomas}
\address{Zoology building \\ Trinity College Dublin \\ Dublin 2, Ireland \\ \\ ncooper@tcd.ie}
\longindentation=0pt
\begin{document}

\begin{letter}{}

Referees' Comments to Authors:
Referee: 1

Comments to the Author
The authors discuss some potential pitfalls in model choice and subsequent interpretation regarding the widely used OU model for analysing a single continuous trait on a phylogenetic tree. Due to the heavy usage of this model (as evidenced by the authors own literature survey) this is clearly an important paper, and the authors provide a broad array of simulations, which span a sufficient number of scenarios to characterise some major biases of the typical likelihood based model-fitting approach. The authors' interpretation of the results is measured and sensible (neither overly doom laden nor free of potential solutions). They have also strived to make their study repeatable whilst acknowledging the challenges in doing so and the limitations of their chosen approach. I have no problem supporting the publication of this paper, although I have a few notes (below) on potential improvements. These overwhelmingly ask for additional clarity on specific points and so I see them as minor.

L22 - Except these are exclusively for continuous data. Maybe use "core" instead of foundation, or a caveat like "at least for continuous data".
L28 (and L199) - I'm not sure if Type 1 error is the correct term to use as this is (to my mind) a frequentists term. I realise that we are (in a sense) rejecting the null of Brownian motion (albeit falsely) and that the LRT gives us a p-value to do so. However, the notion isn't necessarily generalisable to, e.g., three or more models. Perhaps the authors could just state explicitly what they mean by Type I errors to avoid confusion.
L89 - The year range given here and that in the figure/figure caption are different.
L102 - This explanation seems a little incomplete. Why isn't it?
L124 - This also seems incomplete. Maybe insert "merely"?
L156 - I think it would be helpful to discuss mu a bit more. Questions I would like to see answered include: Presumably this is the trait "optimum"? What does it mean to have an optimum outside of the range of the tip values? What if mu is unrealistic, e.g., continuous measure is wingspan and mu is 1000km? If mu is large/far from the tip mean does that mean alpha can be small and the model still be a better fit? etc.
L167 - I'm not totally clear what is meant by tree height (I think I know, but would prefer it to be defined explicitly), nor how or why it scales with alpha.
L188 - Lol.
L219 - Why did the authors chose a Yule model here and a birth-death earlier? Does it mean one is ultrametric and the other not? (Maybe this relates to the fossil point below.)
L263 - This paragraph could do with some unpacking. What asymptotic properties? How is alpha bounded? etc.
L267 - Does this extend, then, to AIC(c)s? (As many people use AIC instead of the LRT and fit more than two models this could be important.)
L283 - "in fitting model" missing "the"?.
L299 - Citations should be in date order.
L325 - Do you mean "regardless OF"?
L330 - Potential pregnancy (missing period).
L342 - "processES"?
L367 - Although in practice are there not ways to set limits on the min-max value of alpha?
L376 - Maybe clarify exactly why (I think I have an idea!).
L382 - Could or should? (Conflict of interest: I am a palaeobiologist!) It seems like fossils could really help as by definition they have smaller values of T, and hence are closer to the root, and hence are better at constraining the root, and hence (potentially) help with model choice in general. (E.g., AFAIK you can only really favour a trend model if you have fossils in the tree.) In addition, when the authors simulate the birth-death models are they using non-ultrametric trees? Or just pruning extinct tips? If the latter then they aren't really testing whether fossils can help, but maybe this could be mentioned as a possible future simulation study.
L385 - Maybe I'm missing something here, but to my mind a non-ultrametric tree is not special in terms of co-variance. Rather it is the variances that are unusual. I.e., the diagonal of the VCV will not just be the same number (same T) for all species. In other words if you modified a non-ultrametric tree to make it ultrametric by making all tips range through to the "present" you would only be modifying the diagonal.
L414 - Epilogue is fine in terms of content, but language is very informal ("vaguely", "messy", "!"). Maybe this isn't an issue, but does read to me as a definite shift in tone from the rest of the paper.
Figure 2 caption - I think this caption could do more to explain the plot. What is meant by "rooty"? What is meant by "tippy"? What does each line represent? What are the dashed lines? etc.
Figure 3 - Would be nice to have more "breaks" in the histogram to help visualise how the data is really distributed. The binning seems too coarse at the moment.

Referee: 2

Comments to the Author
In this article, Cooper and colleagues explore through simulations one of phylogenetic comparative biologists worst kept but most ignored secrets — that the OU model (or, more specifically the single stationary peak version of the OU model) has poor statistical performance. Using a large number of simulated phylogenies derived under different branching processes, they find that: 1) OU has an undesirably high Type I error rate using Likelihood Ratio Tests (LRT), particularly for small trees; 2) ML estimates of the OU α - parameter are erroneously biased towards large values for small trees, and; 3) when measurement error is not accounted for, large datasets, rather than small datasets become more biased towards erroneously favoring OU. The authors provide a suite of practical recommendations for users fitting OU models to comparative datasets and even demonstrate, rather nicely, that using Bayesian model comparison machinery can lead to lower false positive rates than are obtained through ML methods.

Overall, this is a nice contribution that will make a useful reference piece for comparative biologists using PCMs. BJLS is an appropriate outlet - although not normally associated with these kinds of papers, many contributions use these kinds of methods and I’m sure this one will be much read and cited.

However, while I understand the authors’ frustrations with the fitting and interpretations of single peak OU processes that are common in the literature, I think they miss an opportunity to really emphasize their concerns here and demonstrate to readers that it’s easy to correct oneself when support for OU comes back in comparative analyses. Though the authors do note that a best fitting OU model could be nothing more than a noisy BM processes based on parameter estimates, a closer look at their simulation results reveals that this is also the case here. For example, on their simulated trees rescaled to unit height, the highest median alpha estimated for the BM datasets is 0.165 for the 25 tip Yule tree. This translates into a phylogenetic half life of 4.2 time units however, 4 times the age of the simulated clade, which no sane macroevolutionary biologist would interpret as being too different from Brownian motion at this time scale! Furthermore, the upper 95% HPDs (Quantiles surely?) are, in many cases low enough to result in questionable acceptance of OU over BM for any of the simulations.

I think the authors could do a better job of emphasizing the implications of OU model parameters on the expected outcomes of the process with respect to their own simulations, perhaps by adding a section after line 211 on page 9 that explores alpha estimates and their associated phylogenetic half-lives. If the aim here is to caution readers that a best-fitting OU process does not always translate to evolution about a  constrained peak, then showing this with parameter estimates from the simulations would be a useful contribution. This kind of constructive comment would also help those readers who would view this only as a critique of OU models.

As i thought about this, it also occurred to me that it would probably be useful to more completely lay out how the OU process works. I suspect (from my own dealings with authors during reviews) that many do not have an intuitive feel for how the model works still. Based on this, I’d suggest that the authors move their appendix to the main ms and integrate it with the OU model outline on lines 139-174. It may additionally be useful to illustrate the process, both as a diffusion process and as its effects on the phylogenetic variance covariance structure. These are suggestions only, but I suspect that providing a more complete background to the models would further aid naive readers / students who are to be the main beneficiaries of this paper.

Minor comments

SSP vs single peak OU.

I think it would be useful to clarify early on in the ms that the scenario being explored is specifically a single stationary peak model (a la Harmon et al. 2010), rather than a single optimum model. Although I would agree that most users of these methods are more familiar with applying the latter term, there is a small distinction. A single peak OU process may have the root state the same as or different from the long term optimum, whereas the SSP model has both the same. Furthermore, interpretation of these two models with α ≅ 0 is slightly different. For SSP, α ≅ 0 approximates BM, but for a single peak model where the peak differs from the root state, α ≅ 0 collapses to BM with a trend (e.g., see Hansen 1996, Benson et al. 2014). Because the long term mean is introduced as a parameter of the OU process in eq 2 (line 156), it is perhaps worth pointing out this distinction at some point thereafter. Of course, the case where the root state and long term mean are not equal cannot be fitted when fossil data / strong root or internal node priors are lacking and so is unavailable to most PCM users anyway. But still, for completeness I think this is worth pointing out.

line 114. Perhaps my own bugaboo, but I hate the term “to control for phylogeny”. Perhaps replace this with “to model phylogenetically structured residual error in evolutionary correlations”. Revell (2010) might also make a nice citation here.

Line 172: at the risk of losing a citation (although i think you’ll retain me elsewhere) Slater and Pennell used a rate half-life in the context of the exponential decline in rate under the ACDC model, not a phylogenetic half-life for the exponential approach of a trait to its long term mean under the OU model.

The stepping stone analyses and results are very interesting. In Baele et al. (2012), which introduced path sampling for Bayesian demographic model selection using the BEAST software, the authors showed via simulation that Path Sampling and Stepping Stones performed much better than harmonic mean estimators or other model choice metrics at selecting a time homogeneous model over exponential growth (and also at choosing weak exponential growth over constant population size models). There seems a nice parallel here and this suggests that these kinds of approaches should probably be better integrated into comparative methods,.

Lines 263-269: it may be more useful to be explicit here that LRTs assume that the LR statistic is asymptotically chi-square distributed and that this is unlikely the case here (perhaps you could show this using the simulations?). Link this to lines 288-299 to show how to obtain a simulated distribution for the LRT.

lines 404-407: This works the other way too; not incorporating measurement error could lead one to reject a “high phylogenetic signal” model like EB in favor of BM when EB is intact the true model.

Figure 2 legend what are the black and red dashed lines?

Signed: Graham Slater

References

Baele, G., Lemey, P., Bedford, T., Rambaut, A., Suchard, M. A., & Alekseyenko, A. V. (2012). Improving the accuracy of demographic and molecular clock model comparison while accommodating phylogenetic uncertainty. Molecular biology and evolution, 29(9), 2157-2167.

Benson, R. B., Frigot, R. A., Goswami, A., Andres, B., & Butler, R. J. (2014). Competition and constraint drove Cope's rule in the evolution of giant flying reptiles. Nature communications, 5.

Hansen, T. F. (1997). Stabilizing selection and the comparative analysis of adaptation. Evolution, 1341-1351.

Revell, L. J. (2010). Phylogenetic signal and linear regression on species data. Methods in Ecology and Evolution, 1(4), 319-329.

\end{letter}
\end{document}

### 2015-07-27 Manuscript/ohyou.tex
Edits due to reviewer comments

\title{A cautionary note on the use of Ornstein Uhlenbeck models in macroevolutionary studies}
\author{
  Natalie Cooper$^{1,2,*,\dag}$, Gavin H. Thomas$^{3,\dag}$, Chris Venditti$^{4}$\\ Andrew Meade$^{4}$ and Rob P. Freckleton$^{3}$\\
}
\date{}
\affiliation{\noindent{\footnotesize

  $^1$ School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\ 
  $^2$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
  $^3$ Department of Animal and Plant Sciences, University of Sheffield, Sheffield S10 2TN, UK.\\
  $^4$ School of Biological Sciences, University of Reading, Reading, Berkshire, RG6 6BX, UK.\\
  $^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. 
       Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
  $^\dag$These authors contributed equally.
}}

@@

\section{Abstract}
  Phylogenetic comparative methods are increasingly used to give new insights into the dynamics of trait evolution in deep time. 
  The foundation of these methods is a suite of models that attempt to capture evolutionary patterns by extending the Brownian constant variance model. 
  However, the properties of these models are often poorly understood which can lead to the misinterpretation of results.
  Here we focus on one of these models - the Ornstein Uhlenbeck (OU) model.
  We show that the OU model is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, and that many studies fitting this model use datasets that are small and prone to Type I error.   
  We also show that very small amounts of error in datasets can have profound effects on the performance of OU models.
  Our results suggest that simulating fitted models and comparing with empirical results is critical when fitting OU and other extensions of the Brownian model. 
  We conclude by making recommendations for best practice in fitting OU models in phylogenetic comparative analyses, and for interpreting the parameters of the OU model.

@@

\section{Uses of the OU model}
  The popularity of the OU model has grown extensively in recent years (Fig. \ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supporting Information).
  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
  For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}. 
  It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilising selection'', this is inaccurate and misleading. 
  As formulated by \citet{hansen1997stabilizing}, a trait has a primary optimum that is the mean of individual species optimum for that trait. 
  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central value \citep[the primary optimum;][]{hansen2012adaptive}. 
  However, this is not an estimate of stabilising selection in the population genetics sense.

@@

  The OU model is most commonly used to model the evolution of a single continuous character (Table \ref{table.uses}; Supporting Information). 
  Usually several models of evolution \citep[e.g., Brownian motion, OU, Early burst etc.;][]{Harmon:2010aa,cooper2010body,cardillo2015geographic,slater2015iterative} are fit to the same continuous character and model selection is then used to determine which model best fits the data.  
  OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa,beaulieu2012modeling}. 
  The latter represents evolution under multiple selective regimes, and may be more biologically realistic for many datasets. 
  OU models with various numbers of optima are often included in the pool of evolutionary models being compared \citep[e.g.,][]{christin2013anatomical}(Table \ref{table.uses})
  OU models are also commonly used to control for phylogeny in correlative analyses (Table \ref{table.uses}; Supporting Information).  
  Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
  This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data. 
  The $\alpha$ parameter from an OU model can also be used to transform the tree and the variance-covariance matrix.
  This is rarely interpreted as corresponding to any kind of process,
  ; instead it improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}. 
  It is not clear that it was originally intended that the OU model would be used in such a context.
  Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface}.

  The majority of papers use the OU model to model the evolution of a continuous character (Table \ref{table.uses}; Supporting Information) therefore, we focus on this use of the OU model in our simulations below. 
  The principles here also apply to a range of other macroevolutionary models that can be fit to continuous data and compared using model testing procedures (e.g., $\kappa$, $\lambda$, $\delta$, ACDC, Early Burst; \citealp{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa,Harmon:2008aa}).
  Note that we focus on OU models with a single optimum trait value because these are more commonly used (Table \ref{table.uses}; Supporting Information) and easier to simulate.  For discussions on the performance of multiple optima OU models we refer the reader to \citep{beaulieu2012ouwie}.

@@

   The OU model \citep{hansen1997stabilizing,Butler:2004aa} is a random walk where trait values revert back towards some ``optimal'' value with an attraction strength proportional to the parameter $\alpha$. 
    The model has the following form:

@@

    \noindent
    Note that this model has two parameters in addition to those of the Brownian model: $\alpha$ and $\mu$. 
    The parameter $\mu$ is a long-term mean, and it is assumed that species traits evolve around this value. 
    $\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean if they evolve away from it. 
    $\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$. For more details see the Appendix.

    When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent. 
    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present.
    Note that $\alpha$ scales with tree height, so it needs to be interpreted relative to this. 
    Generally the simplest solution is to rescale tree heights to 1 \citep[e.g.,][simulations below]{ives2010phylogenetic}. 
    \citet{ives2010phylogenetic} suggest interpreting $log( \alpha)$ (after rescaling trees to a height of 1) rather than raw $\alpha$. They equate $-log( \alpha) = -4$ as a very low, almost Brownian value, and $-log( \alpha) = 4$ as a very high value. 
    Others \citep[e.g.][]{slater2015iterative,slater2013robust,hansen2012interpreting} prefer to use the ``phylogenetic half-life'': $t_\frac{1}{2}$ (see ``Recommendations for interpreting $\alpha$'' below).

@@

  \subsection{Performance of $\alpha$ and Likelihood ratio tests}
    To determine whether $\alpha$ is biased, and whether Likelihood ratio tests are appropriate for use with the OU model, we estimated $\alpha$ for each of our simulated phylogenies and data, and compared the fit of a Brownian model to that of an OU model using a Likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT (\citealp{Thomas:2011aa}; \href{https://github.com/ghthomas/motmot}{github.com/ghthomas/motmot}). 
    This mirrors the common situation where researchers fit Brownian and OU models and then use Likelihood ratio tests to select the ``best'' model.
    We then estimated the rejection rate of the null (Brownian) model for each set of simulations. 
    This gives us a measure of Type I error rate for the OU model.

      \begin{table}[!htbp]
        \begin{center}
        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model on a range of constant-rate birth death trees. Tree type refers to the extinction fraction for the birth-death trees. The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
        \bigskip
        \input{Tables/OU_bd_noerror.tex}
        \label{tables:bd_noerror}
        \end{center}
\end{table} 
% Weird issue with endfloat means these need to have no white space

      \begin{table}[!htbp]
        \begin{center}
        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model on trees simulated under time variable speciation rates. The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
        \bigskip
        \input{Tables/OU_var_rate_noerror.tex}
        \label{tables:varrate_noerror}
        \end{center}
\end{table} 
% Weird issue with endfloat means these need to have no white space

@@

    \begin{figure}[!htbp]
      \centering
      \includegraphics[width=6cm, height=12cm, keepaspectratio=true]{Figures/OhYou_Figure2.pdf}
      \caption{Examples of profile likelihoods for selected simulated datasets. In all cases the ``true'' value of $\alpha$ is 0.
      }
      \label{figure:ouprofiles}
\end{figure}
% Weird issue with endfloat means these need to have no white space

    \begin{figure}[!htbp]
      \centering
      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure3.pdf}
      \caption{The number of taxa in phylogenies used to fit OU models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014. Two studies with \textgreater 3000 taxa have been omitted for clarity. See Supporting Information for details.
      }
      \label{figure:ntaxa}
\end{figure} 
% Weird issue with endfloat means these need to have no white space

@@

@@

    \begin{landscape}
      \begin{table}[!htbp]
        \begin{center}
        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model with 0, 1, 5, or 10\% measurement error (m.e.). The rejection rate is the proportion of OU models favoured relative to a Brownian motion model.}
        \bigskip
        \input{Tables/OU_yule_error.tex}
        \label{tables:yule_error}
        \end{center}
\end{table} 
% Weird issue with endfloat means these need to have no white space
    \end{landscape}

@@

      \item \textbf{Consider Bayesian approaches.}
      An alternative but less explored approach is to use Bayesian methods. 
      Bayesian model fitting has not been widely available for OU model fitting until recently but is now possible in R packages including diversitree \citep{FitzJohn:2012aa} and GEIGER \citep[the Single Stationary Peak model in \texttt{fitContinuousMCMC};][]{Harmon:2008aa} and stand alone software including BayesTraits \citep{pagel2013bayestraits}.

>>>>

\title{A cautionary note on the use of Ornstein Uhlenbeck models in macroevolutionary studies}
\author{
  Natalie Cooper$^{1,2,3,*,\dag}$, Gavin H. Thomas$^{4,\dag}$, Chris Venditti$^{5}$\\ Andrew Meade$^{5}$ and Rob P. Freckleton$^{4}$\\
}
\date{}
\affiliation{\noindent{\footnotesize

  $^1$ School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\ 
  $^2$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
  $^3$ Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.\\
  $^4$ Department of Animal and Plant Sciences, University of Sheffield, Sheffield S10 2TN, UK.\\
  $^5$ School of Biological Sciences, University of Reading, Reading, Berkshire, RG6 6BX, UK.\\
  $^*$ Corresponding author: nhcooper12@gmail.com, Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.\\ Tel: +44 20 7942 5000.\\
  $^\dag$These authors contributed equally.
}}

@@

\section{Abstract}
  Phylogenetic comparative methods are increasingly used to give new insights into the dynamics of trait evolution in deep time. 
  For continuous traits the core of these methods is a suite of models that attempt to capture evolutionary patterns by extending the Brownian constant variance model. 
  However, the properties of these models are often poorly understood which can lead to the misinterpretation of results.
  Here we focus on one of these models - the Ornstein Uhlenbeck (OU) model.
  We show that the OU model is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, and that many studies fitting this model use datasets that are small and prone to this problem.   
  We also show that very small amounts of error in datasets can have profound effects on the inferences derived from OU models.
  Our results suggest that simulating fitted models and comparing with empirical results is critical when fitting OU and other extensions of the Brownian model. 
  We conclude by making recommendations for best practice in fitting OU models in phylogenetic comparative analyses, and for interpreting the parameters of the OU model.

@@

\section{Uses of the OU model}
  The popularity of the OU model has grown extensively in recent years (Fig. \ref{figure.literature}); even just between 2012 and 2014 over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published (Google Scholar search 15th March 2015; see Supporting Information).
  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
  For example, fit to an OU model is used as evidence for processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}. 
  It is important to note however, that although the OU model is frequently described and interpreted as a model of ``'stabilising selection'', this is inaccurate and misleading. 
  As formulated by \citet{hansen1997stabilizing}, a trait has a primary optimum that is the mean of individual species optimum for that trait. 
  Under this formulation, $\alpha$ can be considered as the strength of the pull towards a central trait value \citep[the primary optimum;][]{hansen2012adaptive}. 
  However, this is not an estimate of stabilising selection in the population genetics sense, where it is a measure of selection within a population towards a fitness optimum on an adaptive landscape (Lande, 1976). This is a qualitatively different process to trait evolution among species which is more akin to a trait tracking movement of the adaptive optima itself.

@@

  The OU model is most commonly used to model the evolution of a single continuous character (Table \ref{table.uses}; Supporting Information). 
  Usually several models of evolution \citep[e.g., Brownian motion, OU, Early burst etc.;][]{Harmon:2010aa,cooper2010body,cardillo2015geographic,slater2015iterative} are fit to the same continuous character and model selection is then used to determine which model best fits the data.  
  OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa,beaulieu2012modeling}. 
  The latter represents evolution under multiple selective regimes, and may be more biologically realistic for many datasets. 
  OU models with various numbers of optima are often included in the pool of evolutionary models being compared \citep[e.g.,][]{christin2013anatomical}(Table \ref{table.uses})
  OU models are also commonly used model phylogenetically structured residual error in evolutionary correlations (\citealp{revell2010phylogenetic}; often colloquially referred to as ``controlling for phylogeny''; Table \ref{table.uses}; Supporting Information).  
  Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
  This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data. 
  The $\alpha$ parameter from an OU model can also be used to transform the tree and the variance-covariance matrix.
  This is rarely interpreted as corresponding to any kind of process,
  ; instead it improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}. 
  It is not clear that it was originally intended that the OU model would merely be used in such a context.
  Finally, the OU model is also used to reconstruct ancestral states \citep{martins1999estimation} and to detect clade-wide convergent evolution \citep{ingram2013surface,uyeda2014novel}.

  The majority of papers use the OU model to model the evolution of a continuous character (Table \ref{table.uses}; Supporting Information) therefore, we focus on this use of the OU model in our simulations below. 
  The principles here also apply to a range of other macroevolutionary models that can be fit to continuous data and compared using model testing procedures (e.g., $\kappa$, $\lambda$, $\delta$, ACDC, Early Burst; \citealp{Pagel:1997aa,Pagel:1999aa,Blomberg:2003aa,Harmon:2008aa}).
  Note that we focus on OU models with a single stationary optimum trait value because these are more commonly used (Table \ref{table.uses}; Supporting Information) and easier to simulate.  For discussions on the performance of multiple optima OU models we refer the reader to \citep{beaulieu2012ouwie}.

@@

   The OU model \citep{hansen1997stabilizing,Butler:2004aa} is a random walk where trait values revert back towards some ``optimal'' value, $\mu$ (also called $\theta$) with an attraction strength proportional to the parameter $\alpha$. 
    The model has the following form:

@@

    \noindent
    Note that this model has two parameters in addition to those of the Brownian model: $\alpha$ and $\mu$. 
    $\alpha$ is the strength of evolutionary force that returns traits back towards the long-term mean, $\mu$, if they evolve away from it.$\alpha$ is sometimes referred to as the ``rubber band'' parameter because of the way it pulls traits back towards $\mu$.
    The parameter $\mu$ is a long-term mean, and it is assumed that species traits evolve around this value. 
    For more details see the Appendix.

    In many implementations of the OU model \citep[e.g. GEIGER][]{Harmon:2008aa}, $\mu$ is the same as the state of the root at time zero, $X(0)$. 
    This is referred to as a ``single stationary peak'' or SSP model (\textit{sensu} \citealp{Harmon:2010aa}). Some implementations also allow users to estimate $X(0)$ \citep[e.g. OUwie;][]{beaulieu2012modeling}. 
    Where $X(0) \neq \mu$ this is referred to as a single peak OU process. 
    However, estimating $X(0)$ can sometimes lead to nonsensical values of $\mu$, i.e. far outside the range of values for the trait, thus results should be interpreted with caution. $X(0)$ can also be defined \textit{a priori}, but this is only appropriate when fossil data, or other independent evidence, allow confident estimates of the root state.

    When $\alpha$ is close to zero, evolution is approximately Brownian, (but note that in the special case of a single peak OU model with $X(0) \neq \mu$, when $\alpha$ is zero, evolution approximates Brownian motion with a trend; \citealp{hansen1997stabilizing,benson2014competition}), then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent. 
    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present.
    Note that $\alpha$ scales with tree height (i.e. the maximum distance from the root of the tree to the tips); taller trees will have lower $\alpha$ values, all else being equal, because there is more time for traits to return to the optimum value thus the strength of the pull towards the optimum, $\alpha$, can be smaller.
    Thus $\alpha$ values need to be interpreted relative to tree height.Generally the simplest solution is to rescale tree heights to 1 \citep[e.g.,][simulations below]{ives2010phylogenetic}. 
    \citet{ives2010phylogenetic} suggest interpreting $log( \alpha)$ (after rescaling trees to a height of 1) rather than raw $\alpha$. They equate $-log( \alpha) = -4$ as a very low, almost Brownian value, and $-log( \alpha) = 4$ as a very high value. 
    Others \citep[e.g.][]{slater2015iterative,slater2013robust,hansen2012interpreting} prefer to use the ``phylogenetic half-life'': $t_\frac{1}{2}$ (see ``Recommendations for interpreting $\alpha$'' below).

@@

  \subsection{Performance of $\alpha$ and Likelihood ratio tests}
    To determine whether $\alpha$ is biased under conditions where it should be zero, and whether Likelihood ratio tests are appropriate for use with the OU model, we estimated $\alpha$ for each of our simulated phylogenies and data, and compared the fit of a Brownian model to that of an OU model using a Likelihood ratio test with 1 degree of freedom with the \texttt{transformPhylo.ML} function in MOTMOT (\citealp{Thomas:2011aa}; \href{https://github.com/ghthomas/motmot}{github.com/ghthomas/motmot}). 
    This mirrors the common situation where researchers fit Brownian and OU models and then use Likelihood ratio tests to select the ``best'' model.
    We then estimated the rejection rate of the null (Brownian) model for each set of simulations. 
    We refer to this measure as the Type I error rate for the OU model.

      \begin{table}[!htbp]
        \begin{center}
        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model on a range of constant-rate birth death trees. Tree type refers to the extinction fraction for the birth-death trees. The rejection rate is the proportion of Ornstein Uhlenbeck models favoured relative to a Brownian motion model.}
        \bigskip
        \input{Tables/OU_bd_noerror.tex}
        \label{tables:bd_noerror}
        \end{center}
\end{table} 
% Weird issue with endfloat means these need to have no white space

      \begin{table}[!htbp]
        \begin{center}
        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model on trees simulated under time variable speciation rates. The rejection rate is the proportion of Ornstein Uhlenbeck models favoured relative to a Brownian motion model.}
        \bigskip
        \input{Tables/OU_var_rate_noerror.tex}
        \label{tables:varrate_noerror}
        \end{center}
\end{table} 
% Weird issue with endfloat means these need to have no white space

@@

    \begin{figure}[!htbp]
      \centering
      \includegraphics[width=6cm, height=12cm, keepaspectratio=true]{Figures/OhYou_Figure2.pdf}
      \caption{Examples of profile likelihoods for selected simulated datasets for tippy (A), rooty (B) and Yule (C) simulated trees. 
      Each solid black line represents one simulated dataset selected at random. 
      Tippy trees are those with branching events distributed disproportionately late in the clade's history (i.e. nearer to the present).
      Rooty trees are those with branching events distributed disproportionately early in the clade's history (i.e. nearer to the root). In all cases the ``true'', value of $\alpha$ is 0 (black dashed line). The red dashed line is... % FIX!!!
      }
      \label{figure:ouprofiles}
\end{figure}
% Weird issue with endfloat means these need to have no white space

    \begin{figure}[!htbp]
      \centering
      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/OhYou_Figure3.pdf}
      \caption{The number of taxa in phylogenies used to fit Ornstein Uhlenbeck models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2014. Two studies with \textgreater 3000 taxa have been omitted for clarity. See Supporting Information for details.
      }
      \label{figure:ntaxa}
\end{figure} 
% Weird issue with endfloat means these need to have no white space

@@

    For small trees the confidence limits on parameter estimates are broad (Figure \ref{figure:ouprofiles}). 
    However, the median is typically low irrespective of tree size (see Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}) and even when the OU model is favoured, scrutiny of the model parameters suggest that the favoured OU model is biologically indistinguishable from Brownian motion. 
    Indeed, examination of the effect of $\alpha$ on the expected covariances among taxa (Figure \ref{figure:ouhistory}) confirms that covariances are barely distinguishable when $\alpha$ \textless 1 for unit height trees. 
    This suggests that, notwithstanding the elevated Type I errors of OU models, their interpretation requires examination of parameters. 
    We return to the issue of parameter interpretation below, specifically with reference to the increasingly commonly used ``phylogenetic half-life'', $t_\frac{1}{2}$.

@@

    \begin{landscape}
      \begin{table}[!htbp]
        \begin{center}
        \caption{Rejection rate and $\alpha$ estimates for data simulated under a constant rate Brownian model with 0, 1, 5, or 10\% measurement error (m.e.). The rejection rate is the proportion of Ornstein Uhlenbeck models favoured relative to a Brownian motion model.}
        \bigskip
        \input{Tables/OU_yule_error.tex}
        \label{tables:yule_error}
        \end{center}
\end{table} 
% Weird issue with endfloat means these need to have no white space
    \end{landscape}

@@

      \item \textbf{Consider Bayesian approaches.}
      An alternative but less explored approach is to use Bayesian methods. 
      Bayesian model fitting has not been widely available for OU model fitting until recently but is now possible in R packages including diversitree \citep{FitzJohn:2012aa} and GEIGER \citep[the Single Stationary Peak model in \texttt{fitContinuousMCMC};][]{Harmon:2008aa} and stand alone software including BayesTraits \citep{pagel2013bayestraits}.
      Below we explore Bayesian methods as a possible remedy to the limitations of fitting OU models in a likelihood framework.
